---
title: "Physical Activity Index"
subtitle: "HUNT 1-4, change in PA"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Information

This script contains the analyses for change in LTPA across H1-H3-H4.

Due to different phrasing of the PA questions in HUNT 2, these analyses **will only consider longitudinal trends from H1-H3-H4**. That is, across four decades, with 10 years between 10 years between H3-H4, 22 years between H1-H3 and 32 years between H1-H3-H4.

## Set environment and load packages

```{r setup}

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
```

```{r}

library(tidyverse)
library(survival)
library(purrr)
library(patchwork)
```

## Longitudinal trends across H1-H3-H4

**Spørsmål til Mats**:

1.  Må man begrense oppfølgingstid når man f.eks. ser på H1-H3?
2.  Nå har jeg filtrert ut på part_dat-variablene, er det standard praksis hos dere?
3.  I f.eks. H1-H3 må man lage en PA-gruppe for hvert individ til hver undersøkelse for å studere endring i FA. Hvordan gjør man det med andre kovariater? Bruker man alkoholforbruk både i H1 og H3, eller bruker man bare den ene?
4. 

### Process data

In order to perform the analyses, we must filter out participants as follows:

-   H3-H4: Participants who have data in H3 and H4.
-   H1-H3: Participants who have data in H1 and H3.
-   H1-H3-H4: Participants who have data in H1, H3 and H4.

These operations are in **data_preparation.qmd**

There functions for cleaning the data is in **functions.qmd**.

#### H1-H3

```{r}

```

#### H1-H3-H3

```{r}


```

#### H3-H4

```{r}


```
