---
title: "Data preparation"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Information

This script is a general preparation for the data. It retrieves the data and does some basic processing. In addition, it distinguished the data from the different surveys and saves them in separate data frames-

## Set environment and load packages

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
```

```{r}

library(tidyverse)
```

## Load data

```{r}

full_data <- read_csv("//fil.nice.ntnu.no/nice/p758/Data master/full_data.csv")
```

## Process data

### Basic processing

Un-capitazing names, change variable types and create date-time objects.

```{r}

# Un-capitalize and clean names

full_cleaned_data <- full_data |> 
  janitor::clean_names()

# Remove initial data frame

rm(full_data)

# Change from chr to fct

full_cleaned_data <- full_cleaned_data |>
  mutate_if(is.character, as.factor)

# Manually change actual chr back to fct again

full_cleaned_data <- full_cleaned_data |> 
  mutate(
    birth_date = as.character(birth_date), 
    end_date_death = as.character(end_date_death),
    end_date_icd8910 = as.character(end_date_icd8910),
    end_date_link = as.character(end_date_link),
    w22_0389_lopenr_person = as.character(w22_0389_lopenr_person),
    death_all = as.factor(death_all),
    death_all_icd = as.factor(death_all_icd),
    death_cvd = as.factor(death_cvd),
    death_ihd = as.factor(death_ihd),
    death_stroke = as.factor(death_stroke),
    death_ihd = as.factor(death_ihd)
    )

# Create date objects

full_cleaned_data <- full_cleaned_data |> 
  mutate(
    part_dat_nt1blq1 = dmy(part_dat_nt1blq1),
    part_dat_nt2blq1 = dmy(part_dat_nt2blq1),
    part_dat_nt3blq1 = dmy(part_dat_nt3blq1),
    part_dat_nt4blq1 = dmy(part_dat_nt4blq1),
    end_date_death = dmy(end_date_death),
    end_date_link = dmy(end_date_link),
    end_date_icd8910 = dmy(end_date_icd8910),
    eof_date_death = dmy(eof_date_death)
  )

# Compute participant age

full_cleaned_data <- full_cleaned_data |> 
  mutate(age = year(end_date_death) - birth_year)
```

### Distinguish different HUNT surveys into separate data frames

#### Data frames for no change in LTPA

Since all the data for all four surveys are in one dataset, we have to distinguish them into four separate data sets.

Each survey has two variables called BLM and a variable related to each sub-survey. We use these to distinguish.

#### HUNT 1

Q2 is the relevant Q for H1, it contains data on exercise, work, education, smoking and alcohol

```{r}

hunt_1_cleaned_full_data <- full_cleaned_data |> 
  select(contains(match = "nt1"), age, sex, end_date_death, death_all) |> 
  filter(!is.na(part_nt1blq2) & !is.na(part_nt1blm))
```

#### HUNT 2

Q1 is the relevant for Q for H2, it contains data on exercise, work, education, smoking and alcohol

```{r}

hunt_2_cleaned_full_data <- full_cleaned_data |> 
  select(contains(match = "nt2"), age, sex, end_date_death, death_all) |> 
  filter(!is.na(part_nt2blq1) & !is.na(part_nt2blm))
```

#### HUNT 3

Q1 is the relevant for Q for H3, it contains data on exercise, work, education, smoking and alcohol

```{r}

hunt_3_cleaned_full_data <- full_cleaned_data |> 
  select(contains(match = "nt3"), age, sex, end_date_death, death_all) |> 
  filter(!is.na(part_nt3blq1) & !is.na(part_nt3blm))
```

#### HUNT 4

Have no information on the surveys, so just did a filtering on baseline measurements (i.e, blm variable)

```{r}

hunt_4_cleaned_full_data <- full_cleaned_data |>
  select(contains(match = "nt4"), age, sex, end_date_death, death_all) |> 
  filter(!is.na(part_nt4blm))
```

#### Data frames for change in LTPA

##### HUNT 3-4

```{r}

hunt_3_4 <- full_cleaned_data |> 
  select(contains(match = c("nt3", "nt4"))) |> 
  filter(
    !is.na(part_nt3blq1) & !is.na(part_nt3blm) & !is.na(part_nt4blm)
    )
```

##### HUNT 1-3-4

```{r}

hunt_1_3_4 <- full_cleaned_data |> 
  select(contains(match = c("nt1", "nt3", "nt4"))) |> 
  filter(
    !is.na(part_nt1blq2) & !is.na(part_nt1blm) & !is.na(part_nt3blq1) & !is.na(part_nt3blq1) & !is.na(part_nt4blm)
  )
```

































