---
title: "HUNT 1-4: No change in PA"
author: "10 year cut-off"
format: html
editor: visual
---

## Information

This script contains the analyses for all four HUNT surveys without considering change in LTPA stratified by socioeconomic class while having a cut-off on follow-up time in 10 years or less. Through the analyses I have used own functions which is necessary for computing, see functions.qmd.

## Set environment and load packages

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(tidyverse)
library(survival)
library(purrr)
```

## HUNT 1 - Survival Analysis

### Process data

```{r}

hunt_1_cleaned_data_10_yrs <- full_cleaned_data |> 
  process_hunt_1_no_change() |>
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |> 
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
    education_level %in% c(
      "7-årig folkeskole eller kortere",
      "9-årig grunnskole", 
      "Real- eller middelskole, grunnskolen 10.år",
      "Framhalds- eller fortsettelsesskole"
      ) ~ "low",
    education_level %in% c(
      "Ett- eller toårig videregående skole",
      "Artium, økonomisk gymnas eller almenfaglig retning i vgs."
      ) ~ "medium",
    education_level %in% c(
      "Høyskole eller universitet, 4 år eller mer", 
      "Høyskole eller universitet, mindre enn 4 år"
      )~ "high")
    )
  ) |> 
  filter(
    follow_up_time_in_years <= 10 # Constrict follow-up time be only in 10 years
  )
```

### Run Cox models

Crude:

```{r}

hunt_1_cox_reg_crude_10_yrs <- run_cox_reg_crude(hunt_1_cleaned_data_10_yrs, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_1_cox_reg_multi_10_yrs <- run_cox_reg_multi(hunt_1_cleaned_data_10_yrs, strata = socioeconomic_class)
```

## HUNT 2 - Survival Analysis

### Process data

```{r}

hunt_2_cleaned_data_10_yrs <- full_cleaned_data |> 
  process_hunt_2_no_change() |> 
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |>
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
    education_level %in% c(
      "Grunnskole 7-10 år, framhaldsskole, folkehøgskole"
      ) ~ "low",
    education_level %in% c(
      "Artium, øk.gymnas, allmennfaglig retning i videregående skole", 
      "Realskole, middelskole, yrkesskole 1-2 årig videregående skole"
      ) ~ "medium",
    education_level %in% c(
      "Høgskole/universitet, 4 år eller mer", 
      "Høgskole/universitet, mindre enn 4 år"
      ) ~ "high")
    )
  ) |> 
  filter(
    follow_up_time_in_years <= 10
    )
```

### Run Cox models

Crude:

```{r}

hunt_2_cox_reg_crude_10_yrs <- run_cox_reg_crude(hunt_2_cleaned_data_10_yrs, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_2_cox_reg_multi_10_yrs <- run_cox_reg_multi(hunt_2_cleaned_data_10_yrs, strata = socioeconomic_class)
```

## HUNT 3 - Survival Analysis

### Process data

```{r}

hunt_3_cleaned_data_10_yrs <- full_cleaned_data |> 
  process_hunt_3_no_change() |> 
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |>
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
    occupational_type %in% c(
      "Yrker uten krav til utdanning", 
      "Yrker innen jordbruk, skogbruk og fiske",
      "Prosess- og maskinoperatører, transportarbeidere mv."
      ) ~ "low",
    occupational_type %in% c(
      "Kontor- og kundeserviceyrker", 
      "Salgs-, service- og omsorgsyrker", 
      "Håndverkere o.l.", 
      "Militære yrker og uoppgitt"
      ) ~ "medium",
    occupational_type %in% c(
      "Administrative ledere og politikere", 
      "Akademiske yrker", 
      "Yrker med kortere høyskole- og universitetsutdanning og teknikere"
      ) ~ "high")
    )
  ) |> 
  filter(
    follow_up_time_in_years <= 10
    )
```

### Run Cox models

Crude:

```{r}

hunt_3_cox_reg_crude_10_yrs <- run_cox_reg_crude(hunt_3_cleaned_data_10_yrs, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_3_cox_reg_multi_10_yrs <- run_cox_reg_multi(hunt_3_cleaned_data_10_yrs, strata = socioeconomic_class)
```

## HUNT 4 - Survival Analysis

### Process data

```{r}

hunt_4_cleaned_data_10_yrs <- full_cleaned_data |> 
  process_hunt_4_no_change() |> 
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |>
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
      educational_level %in% c(
        "Grunnskole", 
        "Fagbrev eller svennebrev"
        ) ~ "low",
      educational_level %in% c(
        "3 år i videregående skole", 
        "1-2årig videregående skole"
        ) ~ "medium",
      educational_level %in% c(
        "Høyskole/universitet, 4 år eller mer", 
        "Høyskole/universitet, mindre enn 4 år"
        ) ~ "high")
    )
  ) |> 
  filter(
    follow_up_time_in_years <= 10
    )
```

### Run Cox models

Crude:

```{r}

hunt_4_cox_reg_crude_10_yrs <- run_cox_reg_crude(hunt_4_cleaned_data_10_yrs, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_4_cox_reg_multi_10_yrs <- run_cox_reg_multi(hunt_4_cleaned_data_10_yrs, strata = socioeconomic_class)
```
