---
title: "HUNT 1-4: No change in PA"
subtitle: "Stratified by socioeconomic class"
format: html
editor: visual
---

## Information

This script contains the analyses for all four HUNT surveys without considering change in LTPA stratified by occupational physical activity. Through the analyses I have used own functions which is necessary for computing, see functions.qmd.

## Set environment and load packages

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
```

```{r}

# Packages

library(tidyverse)
library(survival)
library(purrr)

# Functions, see functions.R

source("functions.R")
```

## HUNT 1 - Survival Analysis

#### Process data

```{r}

hunt_1_cleaned_data_occu_strat <- hunt_1_cleaned_data |>
  rename(
    heavy_physical_work = wor_heav_nt1blq2
    ) |> 
  mutate(
    occupational_pa = as.factor(case_when(
      heavy_physical_work %in% c(
        "Aldri, eller nesten aldri", 
        "Ganske sjelden"
        ) ~ "low",
      heavy_physical_work %in% c(
        "Ja, nesten alltid", 
        "Ganske ofte"
        ) ~ "high")
      )
    )
```

#### Run Cox models

Crude:

```{r}

hunt_1_cox_reg_crude <- run_cox_reg_crude(hunt_1_cleaned_data, strata = NULL)
```

Multi-adjusted:

```{r}

hunt_1_cox_reg_multi <- run_cox_reg_multi(hunt_1_cleaned_data, strata = NULL)
```

## HUNT 2 - Survival Analysis

#### Process data

```{r}

hunt_2_cleaned_data_occu_strat <- hunt_2_cleaned_data |>
  rename(
    heavy_physical_work = wor_heav_nt2blq2
    ) |> 
  mutate(
    occupational_pa = as.factor(case_when(
      heavy_physical_work %in% c(
        "Aldri, eller nesten aldri", 
        "Ganske sjelden"
        ) ~ "low",
      heavy_physical_work %in% c(
        "Ja, nesten alltid", 
        "Ganske ofte"
        ) ~ "high")
      )
    )
```

#### Run Cox models

```{r}

hunt_2_cox_reg_crude <- run_cox_reg_crude(hunt_2_cleaned_data, strata = NULL)
```

```{r}

hunt_2_cox_reg_multi <- run_cox_reg_multi(hunt_2_cleaned_data, strata = NULL)
```

## HUNT 3 - Survival Analysis

#### Process data

```{r}

hunt_3_cleaned_data_occu_strat <- hunt_3_cleaned_data |>
  rename(
    heavy_physical_work = wor_heav_nt3blq2
    ) |> 
  mutate(
    occupational_pa = as.factor(case_when(
      heavy_physical_work %in% c(
        "Aldri, eller nesten aldri", 
        "Ganske sjelden"
        ) ~ "low",
    heavy_physical_work %in% c(
      "Ja, nesten alltid", 
      "Ganske ofte"
      ) ~ "high")
    )
  )
```

#### Run Cox models

Crude:

```{r}

hunt_3_cox_reg_crude <- run_cox_reg_crude(hunt_3_cleaned_data, strata = NULL)
```

Multi-adjusted:

```{r}

hunt_3_cox_reg_multi <- run_cox_reg_multi(hunt_3_cleaned_data, strata = NULL)
```

## HUNT 4 - Survival Analysis

#### Process data

```{r}

hunt_4_cleaned_data_occu_strat <- hunt_4_cleaned_data |>
  rename(
    heavy_physical_work = wor_heav_nt4blq2
    ) |> 
  mutate(
    occupational_pa = as.factor(case_when(
      heavy_physical_work %in% c(
        "Aldri, eller nesten aldri", 
        "Ganske sjelden"
        ) ~ "low",
      heavy_physical_work %in% c(
        "Ja, nesten alltid", 
        "Ganske ofte"
        ) ~ "high")
    )
  )
```

#### Run Cox models

Crude:

```{r}

hunt_4_cox_reg_crude <- run_cox_reg_crude(hunt_4_cleaned_data, strata = NULL)
```

Multi-adjusted:

```{r}

hunt_4_cox_reg_multi <- run_cox_reg_multi(hunt_4_cleaned_data, strata = NULL)
```

## Summary statistics from the models

### Crude model

#### Create vectors with data frames and covariates

```{r}

# Create a vector of the data frames 

dataframes_occu_strat <- c(
  "hunt_1_cleaned_data_occu_strat", 
  "hunt_2_cleaned_data_occu_strat", 
  "hunt_3_cleaned_data_occu_strat",
  "hunt_4_cleaned_data_occu_strat"
)

# Create vector of the covariates for the multi-adjusted model

crude_covariates_occu_strat <- c(
  "follow_up_time_in_years", 
  "death_all_cause", 
  "age", 
  "pa_hrs_per_week"
)

# Create a vector of the occupational PA level

occupational_pa_classes <- c(
  "low", 
  "high"
)
```

#### Compute summary statistics

```{r}

# Number of participants

calculate_num_of_participants(
  dataframes = dataframes,
  covariates = multi_adjusted_covariates
)
```

```{r}

# Person-years follow-up

calculate_person_years_follow_up(
  dataframes = dataframes, 
  covariates = multi_adjusted_covariates
)
```

```{r}

# Number of deaths

calculate_num_of_deaths(
  dataframes = dataframes,
  covariates = multi_adjusted_covariates
)
```

### Multi-adjusted model

#### Create vectors with data frames and covariates

```{r}

# Create vector of the covariates for the multi-adjusted model

multi_adjusted_covariates_occu_strat <- c(
  "age", 
  "pa_hrs_per_week", 
  "follow_up_time_in_years", 
  "death_all_cause", 
  "bp_systolic",
  "bmi", 
  "smo_status", 
  "sex", 
  "alcohol_usage", 
  "occupational_pa"
)

# We're re-using the data frames vector and the occupational PA vector from the crude model
```

```{r}

# Number of participants

calculate_num_of_participants(
  dataframes = dataframes,
  covariates = crude_covariates
)
```

```{r}

# Person-years follow-up

calculate_person_years_follow_up(
  dataframes = dataframes, 
  covariates = crude_covariates
)
```

```{r}

# Number of deaths

calculate_num_of_deaths(
  dataframes = dataframes,
  covariates = crude_covariates
)
```

### Combine all data from the models in a tibble

This is done for readability.

```{r}


tibble(
  survey = c("HUNT 1", "HUNT 2", "HUNT 3", "HUNT 4"),
  number_of_participants = c(33926,50931, 36118, 38430),
  num_of_deaths = c(17011, 13992, 4024, 405),
  person_years_follow_up = c(931806, 1093581, 458556, 102768),
  incidence_rate = (num_of_deaths) / (person_years_follow_up) * 1000,
  multi_hr = c(1.040, 0.946, 0.948, 0.881),
  ci_low = c(1.030, 0.933, 0.928, 0.822),
  ci_high = c(1.050, 0.958, 0.968, 0.945)
)

# Crude model

tibble(
  survey = c("HUNT 1", "HUNT 2", "HUNT 3", "HUNT 4"),
  number_of_participants = c(35080, 54397, 38456, 44627),
  num_of_deaths = c(17984, 15362, 4705, 604),
  person_years_follow_up = c(953010, 1160780, 485785, 119545),
  incidence_rate = (num_of_deaths) / (person_years_follow_up) * 1000,
  crude_hr = c(1.070, 0.920, 0.951, 0.846),
  ci_low = c(1.060, 0.909, 0.932, 0.797),
  ci_high = c(1.080, 0.932, 0.970, 0.898)
)
```
