---
title: "HUNT 1-4: No change in PA"
author: "3 year delay for registration"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Information

This script contains the analyses for all four HUNT surveys without considering change in LTPA stratified by socioeconomic class while having a delay on follow-up in 3 years. Naturally, due to the short follow-up time in HUNT 4, that analysis is excluded.

Through the analyses I have used own functions which is necessary for computing, see functions.qmd.

## Set environment and load packages

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

```{r}

library(tidyverse)
library(survival)
```

## HUNT 1 - Survival Analysis

### Process data

```{r}

hunt_1_cleaned_data_3_yrs_delay <- full_cleaned_data |> 
  process_hunt_1_no_change() |>
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |> 
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
    education_level %in% c(
      "7-årig folkeskole eller kortere",
      "9-årig grunnskole", 
      "Real- eller middelskole, grunnskolen 10.år",
      "Framhalds- eller fortsettelsesskole"
      ) ~ "low",
    education_level %in% c(
      "Ett- eller toårig videregående skole",
      "Artium, økonomisk gymnas eller almenfaglig retning i vgs."
      ) ~ "medium",
    education_level %in% c(
      "Høyskole eller universitet, 4 år eller mer", 
      "Høyskole eller universitet, mindre enn 4 år"
      )~ "high")
    )
  ) |> 
  filter(
    follow_up_time_in_years >= 3
  )
```

### Run Cox models

Crude:

```{r}

hunt_1_cox_reg_crude_3_yrs_delay <- run_cox_reg_crude(hunt_1_cleaned_data_3_yrs_delay, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_1_cox_reg_crude_3_yrs_delay <- run_cox_reg_multi(hunt_1_cleaned_data_3_yrs_delay, strata = socioeconomic_class)
```

## HUNT 2 - Survival Analysis

### Process data

```{r}

hunt_2_cleaned_data_3_yrs_delay <- full_cleaned_data |> 
  process_hunt_2_no_change() |> 
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |>
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
    education_level %in% c(
      "Grunnskole 7-10 år, framhaldsskole, folkehøgskole"
      ) ~ "low",
    education_level %in% c(
      "Artium, øk.gymnas, allmennfaglig retning i videregående skole", 
      "Realskole, middelskole, yrkesskole 1-2 årig videregående skole"
      ) ~ "medium",
    education_level %in% c(
      "Høgskole/universitet, 4 år eller mer", 
      "Høgskole/universitet, mindre enn 4 år"
      ) ~ "high")
    )
  ) |> 
  filter(
    follow_up_time_in_years >= 3
  )
```

### Run Cox models

Crude:

```{r}

hunt_2_cox_reg_crude_3_yrs_delay <- run_cox_reg_crude(hunt_2_cleaned_data_3_yrs_delay, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_2_cox_reg_crude_3_yrs_delay <- run_cox_reg_multi(hunt_2_cleaned_data_3_yrs_delay, strata = socioeconomic_class)
```

## HUNT 3 - Survival Analysis

### Process data

```{r}

hunt_3_cleaned_data_3_yrs_delay <- full_cleaned_data |> 
  process_hunt_3_no_change() |> 
  select(
    -c(heavy_physical_work) # Not necessary for computation, so it only limits statistical power. Therefore, removed
  ) |>
  drop_na() |> # All NA's dropped because the same participants has to be included in the crude and multi-adjusted model
  mutate(
    socioeconomic_class = as.factor(case_when(
    occupational_type %in% c(
      "Yrker uten krav til utdanning", 
      "Yrker innen jordbruk, skogbruk og fiske",
      "Prosess- og maskinoperatører, transportarbeidere mv."
      ) ~ "low",
    occupational_type %in% c(
      "Kontor- og kundeserviceyrker", 
      "Salgs-, service- og omsorgsyrker", 
      "Håndverkere o.l.", 
      "Militære yrker og uoppgitt"
      ) ~ "medium",
    occupational_type %in% c(
      "Administrative ledere og politikere", 
      "Akademiske yrker", 
      "Yrker med kortere høyskole- og universitetsutdanning og teknikere"
      ) ~ "high")
    )
  ) |>
  filter(
    follow_up_time_in_years >= 3
  )
```

### Run Cox models

Crude:

```{r}

hunt_3_cox_reg_crude_3_yrs_delay <- run_cox_reg_crude(hunt_3_cleaned_data_3_yrs_delay, strata = socioeconomic_class)
```

Multi-adjusted:

```{r}

hunt_3_cox_reg_crude_3_yrs_delay <- run_cox_reg_multi(hunt_3_cleaned_data_3_yrs_delay, strata = socioeconomic_class)
```







