---
title: "No change in PA"
author: "Simen Løkken"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Information

This script contains the analyses for HUNT 1-4 without considering change in LTPA.

## Set environment and load packages

```{r setup}

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
```

```{r}

library(dplyr)
library(tibble)
library(lubridate)
library(readr)
library(tidyr)
library(arrow)
library(forcats)
library(purrr)
library(survival)
```

## Association between LTPA and risk of all-cause mortality

In order to perform the analyses, we must filter out participants from each survey. This is done in data_preparation.qmd.

Then, we must a data processing function to each data set, these functions lie in functions.qmd.

**Note that we re-use the vectors containing the crude and multi-adjusted covariates throughout all the models.**

### H1

#### No stratification

##### Process data

Read in data:

```{r}

hunt_1 <- read_parquet(
  "//fil.nice.ntnu.no/nice/p758/data/my_data/hunt_1.parquet"
)
```

Apply data processing function:

```{r}

glimpse(hunt_1)

hunt_1 <- hunt_1 |> 
  process_hunt_1()

```

##### Crude Cox Regression

```{r}

crude_covariates <- c("sex", "age", "pa_min_per_week")

run_cox_reg(
  data = hunt_1_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

multi_adj_covariates <- c(
  "pa_min_per_week",
  "bp_systolic",
  "bmi",
  "smo_status",
  "age",
  "sex",
  "alcohol_usage"
)

run_cox_reg(
  data = hunt_1_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

```{r}

```

#### Stratified by socioeconomic status

Add a socioeconomic variable to the data:

```{r}

hunt_1_cleaned_full_data_socio_strat <-  hunt_1_cleaned_full_data |> 
  mutate(
    socioeconomic_class = as.factor(case_when(
      education_level %in% c(
        "7-årig folkeskole eller kortere", 
        "9-årig grunnskole", 
        "Real- eller middelskole, grunnskolen 10.år", 
        "Framhalds- eller fortsettelsesskole") ~ "low",
      education_level %in% c(
        "Ett- eller toårig videregående skole", 
        "Artium, økonomisk gymnas eller almenfaglig retning i vgs.") ~ "medium",
      education_level %in% c(
        "Høyskole eller universitet, 4 år eller mer", 
        "Høyskole eller universitet, mindre enn 4 år") ~ "high")
    )
  )
```

##### Crude Cox regression

```{r}

run_cox_reg(
  data = hunt_1_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_1_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

### H2

#### No stratification

##### Process data

Read in data:

```{r}

hunt_2 <- read_parquet(
  "//fil.nice.ntnu.no/nice/p758/data/my_data/hunt_2.parquet"
)
```

Apply data processing function:

```{r}

glimpse(hunt_2)

hunt_2 <- hunt_2 |> 
  process_hunt_2()
```

##### Crude Cox Regression

```{r}

run_cox_reg(
  data = hunt_2_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_2_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

```{r}

```

#### Stratified by socioeconomic status

Add a socioeconomic variable to the data:

```{r}

hunt_2_cleaned_full_data_socio_strat <-  hunt_2_cleaned_full_data |> 
  mutate(
    socioeconomic_class = as.factor(case_when(
      education_level %in% c(
        "Grunnskole 7-10 år, framhaldsskole, folkehøgskole"
        ) ~ "low",
      education_level %in% c(
        "Artium, øk.gymnas, allmennfaglig retning i videregående skole", 
        "Realskole, middelskole, yrkesskole 1-2 årig videregående skole"
        ) ~ "medium",
    education_level %in% c(
      "Høgskole/universitet, 4 år eller mer", 
      "Høgskole/universitet, mindre enn 4 år"
      ) ~ "high"))
  )
```

##### Crude Cox regression

```{r}

run_cox_reg(
  data = hunt_2_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_2_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

### H3

#### No stratification

##### Process data

Read in data:

```{r}

hunt_3 <- read_parquet(
  "//fil.nice.ntnu.no/nice/p758/data/my_data/hunt_3.parquet"
)
```

Apply data processing function:

```{r}

glimpse(hunt_3)

hunt_3 <- hunt_3 |> 
  process_hunt_3()
```

##### Crude Cox Regression

```{r}

run_cox_reg(
  data = hunt_3_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_3_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

```{r}

```

#### Stratified by socioeconomic status

Add a socioeconomic variable to the data:

```{r}

hunt_3_cleaned_full_data_socio_strat <-  hunt_3_cleaned_full_data |> 
  mutate(
    socioeconomic_class = as.factor(case_when(
      occupational_type %in% c(
        "Yrker uten krav til utdanning",
        "Yrker innen jordbruk, skogbruk og fiske",
        "Prosess- og maskinoperatører, transportarbeidere mv."
        ) ~ "low",
      occupational_type %in% c(
        "Kontor- og kundeserviceyrker",
        "Salgs-, service- og omsorgsyrker", 
        "Håndverkere o.l.", "Militære yrker og uoppgitt"
        ) ~ "medium",
      occupational_type %in% c(
        "Administrative ledere og politikere", 
        "Akademiske yrker", 
        "Yrker med kortere høyskole- og universitetsutdanning og teknikere"
        ) ~ "high")
    )
  )
```

##### Crude Cox regression

```{r}

run_cox_reg(
  data = hunt_3_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_3_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

### H4

#### No stratification

##### Process data

Read in data:

```{r}

hunt_4 <- read_parquet(
  "//fil.nice.ntnu.no/nice/p758/data/my_data/hunt_4.parquet"
)
```

Apply data processing function:

```{r}

glimpse(hunt_4)

hunt_4 <- hunt_4 |> 
  process_hunt_4()
```

##### Crude Cox Regression

```{r}

run_cox_reg(
  data = hunt_4_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_4_cleaned_full_data,
  time = follow_up_time_in_years,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

```{r}

```

#### Stratified by socioeconomic status

Add a socioeconomic variable to the data:

```{r}

hunt_4_cleaned_full_data_socio_strat <-  hunt_4_cleaned_full_data |> 
  mutate(
    socioeconomic_class = as.factor(case_when(
      education_level %in% c(
        "Grunnskole", "Fagbrev eller svennebrev"
        ) ~ "low",
      education_level %in% c(
        "3 år i videregående skole", 
        "1-2årig videregående skole"
        ) ~ "medium",
      education_level %in% c(
        "Høyskole/universitet, 4 år eller mer", 
        "Høyskole/universitet, mindre enn 4 år"
        ) ~ "high")
    )
  )
```

##### Crude Cox regression

```{r}

run_cox_reg(
  data = hunt_4_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = crude_covariates
)
```

##### Multi-adjusted Cox regression

```{r}

run_cox_reg(
  data = hunt_4_cleaned_full_data_socio_strat,
  time = follow_up_time_in_years,
  strata = socioeconomic_class,
  status = death_all_cause,
  covariates = multi_adj_covariates
)
```

##### Regression results and statistics

#### Copyied from old script

This is just copied in, has not been positioned yet:

# SUMMARY STATS FROM MODELS

# Multi-adjusted model

```{r}

socioeconomic_class <- c("low", "medium", "high")

for (class in socioeconomic_class) {
    calculate_num_of_participants_strat(
      dataframes = c("hunt_1_cleaned_data_socio_strat", "hunt_2_cleaned_data_socio_strat", "hunt_3_cleaned_data_socio_strat", "hunt_4_cleaned_data_socio_strat"), 
      covariates = c("age", "pa_hrs_per_week", "follow_up_time_in_years", "death_all_cause", "bp_diastolic", "bp_systolic",
                     "bmi", "packs_of_smoke_per_year", "sex", "alcohol_usage", "heart_infarction", "socioeconomic_class"),
      strata = class
    )
}

socioeconomic_class <- c("low", "medium", "high")

for (class in socioeconomic_class) {
    calculate_num_of_participants_strat(
      dataframes = c("hunt_1_cleaned_data_socio_strat", "hunt_2_cleaned_data_socio_strat", "hunt_3_cleaned_data_socio_strat", "hunt_4_cleaned_data_socio_strat"), 
      covariates = c("age", "pa_hrs_per_week", "follow_up_time_in_years", "death_all_cause", "bp_diastolic", "bp_systolic",
                     "bmi", "packs_of_smoke_per_year", "sex", "alcohol_usage", "heart_infarction", "socioeconomic_class"),
      strata = class
    )
}

for (class in socioeconomic_class) {
  calculate_person_years_follow_up_strat(
    dataframes = c("hunt_1_cleaned_data_socio_strat", "hunt_2_cleaned_data_socio_strat", "hunt_3_cleaned_data_socio_strat", "hunt_4_cleaned_data_socio_strat"),
    covariates = c("age", "pa_hrs_per_week", "follow_up_time_in_years", "death_all_cause", "bp_diastolic", "bp_systolic",
                   "bmi", "packs_of_smoke_per_year", "sex", "alcohol_usage", "heart_infarction", "socioeconomic_class"),
    strata = class
  )
}

# Crude model

for (class in socioeconomic_class) {
  calculate_num_of_participants_strat(
    dataframes = c("hunt_1_cleaned_data_socio_strat", "hunt_2_cleaned_data_socio_strat", "hunt_3_cleaned_data_socio_strat", "hunt_4_cleaned_data_socio_strat"), 
    covariates = c("pa_hrs_per_week", "age", "follow_up_time_in_years", "death_all_cause", "socioeconomic_class"),
    strata = class
  )
}

for (class in socioeconomic_class) {
  calculate_person_years_follow_up_strat(
    dataframes = c("hunt_1_cleaned_data_socio_strat", "hunt_2_cleaned_data_socio_strat", "hunt_3_cleaned_data_socio_strat", "hunt_4_cleaned_data_socio_strat"),
    covariates = c("pa_hrs_per_week", "age", "follow_up_time_in_years", "death_all_cause", "socioeconomic_class"),
    strata = class
  )
}

```
